---
title: "ANA 515 Assignment 2"
author: "Abhijeet"
date: "18 June 2022"
output: word_document
---

```{r setup, include=FALSE}
#code chunk to call the installed packages
library(dplyr)
library(knitr)
```

## Section 1

The dataset is a combination of hate crimes data from FBI and Southern Poverty Lawa Center. The data was collected on the basis of key socioeconomic factors for each state, including education, diversity, geographic heterogeneity, economic health, political ideology. The data is measuring the gate crimes per 100k population, and the other columns take into consideration the socioeconomic factors affecting this target variables. 

The research questions that can be asked with the data are 
1) Why are the hate crimes not uniformly distributed across the United States? 
2) What factors influnec the hate crime numbers to rise? 
3) Does the poiltical ideology affect the amount of hate crimes in a specific state? 
4) Do results vary based on the data collected before and after elections? 

The file format is csv. Yes, it is a delimited file which uses comma to seperate the values. 

## Section 2

```{r, include=TRUE}
#using read.csv to read data from csv file from a URL
hate_crimes <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/hate-crimes/hate_crimes.csv")
```

## Section 3
```{r}
#Removing share_population_with_high_school_degree since the it is amost same for all the rows and the column data is from 2009 which does not match the year rest of data was collected
subset1 = select(hate_crimes,-share_population_with_high_school_degree)

#Sorts the states by Median household income
sort_subset1 = arrange(subset1,median_household_income)

hate_crimes  = sort_subset1
```


## Section 4

This dataframe has `r nrow(hate_crimes)` rows and `r ncol(hate_crimes)` columns. The names of the columns and a brief description of each are in the table below:

```{r results = TRUE}
#this makes a new data.frame called text_tbl with two columns, Names and Description
df_table <- data.frame (Name  = c("state", "median_household_income","share_unemployed_seasonal","share_population_in_metro_areas","share_non_citizen","share_white_poverty","gini_index","share_non_white","share_voters_voted_trump","hate_crimes_per_100k_splc","avg_hatecrimes_per_100k_fbi"),
                        Description = c("State","Median household income, 2016", "Share of the population that is unemployed (seasonally adjusted), Sept. 2016","Share of the population that lives in metropolitan areas, 2015","Share of the population that are not U.S. citizens, 2015","Share of white residents who are living in poverty, 2015
                                        ","Gini Index, 2015","Share of the population that is not white, 2015","Share of 2016 U.S. presidential voters who voted for Donald Trump","Hate crimes per 100,000 population, Southern Poverty Law Center, Nov. 9-18, 2016","Average annual hate crimes per 100,000 population, FBI, 2010-2015")
                        )
kable(df_table, caption = "Columns Description")  #prints the table
```

## Section 5
```{r}
hc_subset = select(hate_crimes,"median_household_income","share_unemployed_seasonal","share_population_in_metro_areas")
```

```{r, include = TRUE}
subset_Summary <-summary(hc_subset) #creates the summary
subset_Summary #prints the summary in your output
```



